Abstraction and Invariance for Algebraically Indexed Types

### 2012-07-10 paper/section-instantiations.tex
Fix typesetting of absolute value operation
Make geometry section more consistent with rest of paper

\section{Geometric Consequences of Abstraction}
\label{sec:instantiations}
We now instantiate our general framework %of the previous section 
with
the indexing theory of \autoref{sec:motivating-examples}, and present
more general and formally-justified theorems for free, type
isomorphisms, and non-definability results.

@@

We define the ``standard'' model of the indexing theory for the
two-dimensional geometry example as follows. Each of the sorts is
interpreted just as its semantic counterpart: $\semSort{\SynTransl{2}}
= \Transl{2}$, $ \semSort{\SynGL{2}} = \GL{2}$, $
\semSort{\SynOrth{2}} = \Orth{2}$, and $ \semSort{\SynGL{1}} =
\GL{1}$.
%  \begin{displaymath}
%    \begin{array}{@{}llll}
%      \semSort{\SynTransl{2}} = \Transl{2}
%      &
%      \semSort{\SynGL{2}} = \GL{2}
%      &
%      \semSort{\SynOrth{2}} = \Orth{2}
%      &
%      \semSort{\SynGL{1}} = \GL{1}
%    \end{array}
%  \end{displaymath}
  Each of the index operations (e.g.,~the group structure and
  determinant) is interpreted by the intended semantic operation, and
  clearly satisfy the axioms in \exref{ex:two-dim-geo-axioms}. The
  assignment of relations for the primitive types $\tyPrimNm{vec}$ and
  $\tyPrimNm{real}$ is given by: % the following definitions:
  \begin{displaymath}
    \begin{array}{ll}
      R^{\bullet}_{\tyPrimNm{vec}} = \{( \vec{0}, \vec{0} ) \}&
      R_{\tyPrimNm{vec}}(B,\vec{t}) = \{ (\vec{v}, B\vec{v} + \vec{t}) \sepbar \vec{v} \in \mathbb{R}^2 \} \\
      R^{\bullet}_{\tyPrimNm{real}} = \{(0,0)\} &
      R_{\tyPrimNm{real}}(k) = \{ (x, kx) \sepbar x \in \mathbb{R} \}
    \end{array}
  \end{displaymath}
  The definitions of $R_{\tyPrimNm{vec}}$ and $R_{\tyPrimNm{real}}$
  %are as expected. % given the informal definitions we gave in
formalise the informal ones from~\autoref{sec:motivating-examples}, and 
  the definitions of %We have set
  $R^\bullet_{\tyPrimNm{vec}}$ and $R^\bullet_{\tyPrimNm{real}}$ 
  %to relate $\vec{0}$ and $0$ to themselves to 
  account for the
  polymorphic zero constants we have assumed.
%\end{example}

%\begin{lemma}
%  The thing about $\Gamma_\Geom$.
%\end{lemma}

\subsection{Theorems for Free}
\label{sec:theorems-for-free}

Consider the type of the area %of a triangle 
function from
\exref{ex:area-of-triangle-1}:
\begin{displaymath}
  \begin{array}{@{}l}
    \mathrm{area} : \forall B\mathord:\SynGL{2}, t\mathord:\SynTransl{2}.\ \\
    \hspace{0.8cm}\tyPrim{vec}{B, t} \to \tyPrim{vec}{B, t} \to \tyPrim{vec}{B, t} \to \tyPrim{real}{|\det B|}
  \end{array}
\end{displaymath}
By \thmref{thm:abstraction}, we can derive the following free theorem. For all $B \in \GL{2}$, $\vec{t} \in \Transl{2}$, and $\vec{x}, \vec{y}, \vec{z} \in \mathbb{R}^2$, we have 
\begin{displaymath}
  |\det B|(\tmSem{\mathrm{area}}\ \vec{x}\ \vec{y}\ \vec{z}) = \tmSem{\mathrm{area}}\ (B\vec{x} + \vec{t})\ (B\vec{y} + \vec{t})\ (B\vec{z} + \vec{t})
\end{displaymath}
Thus, directly from the type of the $\mathrm{area}$ function, we can
see that its index-erasure semantics is (a) invariant under
translations, and (b) if the inputs are subjected to a linear
transformation $B$, the output varies with the absolute value of the
determinant of $B$.

@@

\paragraph{Translations} 
Consider first the group $\Transl{2}$ of translations from
\autoref{sec:type-system-geom-intro} and 
\autoref{sec:affine-vector-ops}. For any quantifier-free type $A$
we have: % the isomorphism:
\begin{eqnarray*}
\forall t\mathord:\SynTransl{2}.\tyPrim{vec}{t}\to A
& \cong &
A[0/t]
\end{eqnarray*}
where $A[0/t]$ denotes substitution of the identity translation $0$
for index variable $t$ in type $A$. This isomorphism formalises the
equivalence between \emph{coordinate-free} geometry, expressed by a
polymorphic type whose first argument can be thought of as the
\emph{origin} with respect to which $A$ is specified, and a
\emph{coordinate-based} geometry, expressed by the type $A[0/t]$, in
which the origin is fixed at $(0,0)$. It is sometimes said that an
affine space is a vector space that's forgotten its origin; we have
captured this in a type isomorphism.

\exref{ex:type-iso} is a special case of this isomorphism; another instance is the
type of vector addition:
\begin{eqnarray*}
\forall t_1\mathord:\SynTransl{2}.\forall t_2\mathord:\SynTransl{2}.\tyPrim{vec}{t_1}\to\tyPrim{vec}{t_2}\to\tyPrim{vec}{t_1+t_2}\\
\cong\quad
\forall t_1\mathord:\SynTransl{2}.\tyPrim{vec}{t_1}\to\tyPrim{vec}{t_1}\quad\cong\quad
\tyPrim{vec}{0}
\end{eqnarray*}
\noindent
\newcommand{\transup}[1]{\mathord\uparrow_{#1}}
\newcommand{\transdn}[1]{\mathord\downarrow_{#1}}
The maps associated with the isomorphism are 
$I = \lambda x. x\ [0]\ 0$ and $J = \lambda y. \Lambda t.\lambda v. \transup A (y)$,
%\begin{eqnarray*}
%I = \lambda x. x\ [0]\ 0 & \text{ and } &
%J = \lambda y. \Lambda t.\ \lambda v. \trans A v t(y)
%\end{eqnarray*}
where terms $\transdn A : A\to A[0/t]$ and $\transup A : A[0/t] \to A$ are defined by induction over the syntax of $A$ (cases omitted for $\transup A$ have definition symmetric to $\transdn A$):
\begin{align*}
\transdn{1} &= \lambda x. x \qquad
\transdn{A_1\times A_2} = \lambda p. (\transdn{A_1}(\pi_1 p),\transdn{A_2}(\pi_2 p))\\ 
\transdn{A_1 + A_2} &= \lambda x.
\textrm{case}\ x\ \textrm{of}\ \textrm{inl}\ y.\textrm{inl}(\transdn{A_1}y); \textrm{inr}\ y.\textrm{inr}(\transdn{A_2}y)\\ 
\transdn{A_1\to A_2} &= \lambda f. \lambda x. \transdn{A_2} (f(\transup{A_1}x))\\ 
\transdn{\tyPrim{vec}{e}} &= \lambda x. x - kv \qquad
\transup{\tyPrim{vec}{e}} = \lambda x. x + kv\\ 
\text{ if }e & \equiv k t + e' \text{ for
  some }k\in\setOfIntegers \text{ and }e'\text{ with }t\notin e'
\end{align*}
To show $I(J(y))$ is contextually equivalent to $y$ is an easy
induction on $A$ using the fact that the vector $0$
is an identity for vector addition. To show $J(I(x))$ is contextually
equivalent to $x$ we must apply the abstraction theorem and use
the fact that $\tmSem{\transup A}$ and $\tmSem{\transdn A}$ 
considered as relations are simply instances of $\rsem A{}{}$.
%\[
%$\rsem A {t \mapsto v} {} = \{ (x, \tmSem{\transup A (x)}) \mid x \in
%\tySem A\}$.
%\]

@@

\subsection{Indefinability}
\label{sec:types-indexed-abelian-groups-indef}

\newcommand{\Grp}{\mathit{Grp}}
For proving non-definability results, we
consider the general case of a sort $\mathsf{G}$ with the equational theory of abelian groups.
We assume a group-indexed primitive type $\tyPrimNm{G}$ and
set of primitive operations typed by $\Gamma_\Grp$ that reflect the group structure at the value level,
with identity $1:\tyPrim{G}{1}$, operation
$(\cdot): \forall a,b\mathord:\mathsf{G}.\ \tyPrim{G}{a} \to \tyPrim{G}{b} \to \tyPrim{G}{a\cdot b}$
and inverse $\ ^{-1}: \forall a\mathord:\mathsf{G}.\ \tyPrim{G}{a} \to \tyPrim{G}{a^{-1}}$.
One instance is the sort~$\SynTransl{2}$ with $0$, $(+)$ and $\mathrm{negate}$ operations.

For index-erasure interpretation, we assume a (not necessarily
abelian) group $(G, 1, \cdot, \ ^{-1})$. Let $Z(G)$ be the centre of
the $G$ %: i.e.,~the set of elements that commute with every other
%element in $G$. We set 
and $\tyPrimSem{G} = G$. The primitive operations
in $\Gamma_\Grp$ are interpreted just as the unit, multiplication and
inverse operations of $G$. For each index context $\Delta$, the set
$\relEnv{E}_\Grp(\Delta)$ of relational environments is defined to be
$\{ \rho_{(H,h)} \sepbar H\textrm{ is a subgroup of }(\Delta
\Rightarrow \mathsf{G})/\equiv, h\textrm{ is a group homomorphism }H
\to Z(G) \}$, where
\begin{displaymath}
  \rho_{(H,h)}\ \tyPrimNm{G}\ (e) = \left\{
    \begin{array}{ll}
      \{ \} & \textrm{if}\ e\not\in H \\
      \{ (x, h(e)\oplus x) \sepbar x \in G \} & \textrm{if}\ e\in H
    \end{array}
  \right.
\end{displaymath}

\begin{lemma}
  For all $\Delta$ and $\rho \in \relEnv{E}_\Grp(\Delta)$,
  $(\eta_\Grp,\eta_\Grp) \in \rsem{\Gamma_\Grp}{\relEnv{E}}\rho$.
\end{lemma}

Writing $i^n$ for $n$-fold iteration of the group operation, we have:
\begin{theorem}
  Let $m$ and $n$ be natural numbers. For all $n$-by-$m$-matrices of
  natural numbers $A$ and $m$-by-$1$ matricies of integers $b$,
  there exists a program $M$ such that
  \begin{equation}
    \label{eq:fo-group-type}
    \Gamma_\Grp \vdash M : \forall i_1,\dots,i_m\mathord:\mathsf{G}.\ \tyPrim{G}{e_1} \to \dots \to \tyPrim{G}{e_n} \to \tyPrim{G}{e}
  \end{equation}
  where $e_j = i_1^{A_{j1}}\cdots i_m^{A_{jm}}$ and $e = i_1^{b_1}\cdots i_m^{b_m}$, if and only if there is a solution $\vec{x}$
  in integers to the equation $A \vec x = b$.
\end{theorem}
The uninhabited type  $\forall t \mathord:
  \SynTransl{2}.\ \tyPrim{vec}{t + t} \to \tyPrim{vec}{t}$
from \exref{ex:uninhabited-type} is an instance of this theorem.
For types indexed by the scaling group $\SynGL{1}$ the situation is more subtle, as
the index-erasure interpretation of $\tyPrim{real}{s}$ is group $\GL{1}$ together with zero.
We can show, nonetheless, that $\forall s \mathord:
  \SynGL{1}.\ \tyPrim{real}{s \cdot s} \to \tyPrim{real}{s}\cong 1$, reflecting 
the fact that any term of this type must be $\lambda x.0$.

>>>>

\section{Geometric Consequences of Abstraction}
\label{sec:instantiations}
We now instantiate our general framework %of the previous section 
with
the indexing theory of \autoref{sec:motivating-examples}, and present
more general and formally-justified free theorems, type
isomorphisms, and non-definability results.

@@

We define the model of the indexing theory for the two-dimensional
geometry instantiation as follows. Each of the sorts is interpreted just as
its semantic counterpart: $\semSort{\SynTransl{2}} = \Transl{2}$, $
\semSort{\SynGL{2}} = \GL{2}$, $ \semSort{\SynOrth{2}} = \Orth{2}$,
and $ \semSort{\SynGL{1}} = \GL{1}$.
%  \begin{displaymath}
%    \begin{array}{@{}llll}
%      \semSort{\SynTransl{2}} = \Transl{2}
%      &
%      \semSort{\SynGL{2}} = \GL{2}
%      &
%      \semSort{\SynOrth{2}} = \Orth{2}
%      &
%      \semSort{\SynGL{1}} = \GL{1}
%    \end{array}
%  \end{displaymath}
Each of the index operations (e.g.,~the group structure and
determinant) is interpreted by the intended semantic operation, and
clearly satisfy the axioms in \exref{ex:two-dim-geo-axioms}. For the
relational interpretations of $\tyPrim{vec}{B,t}$ and
$\tyPrim{real}{s}$, we use the compositional construction from
\autoref{sec:constr-rel-env}, with $R_{\tyPrimNm{vec}}(B,\vec{t}) = \{
(\vec{v}, B\vec{v} + \vec{t}) \sepbar \vec{v} \in \mathbb{R}^2 \}$ and
$R_{\tyPrimNm{real}}(k) = \{ (x, kx) \sepbar x \in \mathbb{R} \}$. The
types of the primitive vector operations presented in
\autoref{sec:motivating-examples} are collected into a context
$\Gamma_\Geom$ with a straightforward interpretation $\eta_\Geom \in
\ctxtSem{\Gamma_\Geom}$.
  % \begin{displaymath}
  %   \begin{array}{l}
  %     R_{\tyPrimNm{vec}}(B,\vec{t}) = \{ (\vec{v}, B\vec{v} + \vec{t}) \sepbar \vec{v} \in \mathbb{R}^2 \} \\
  %     R_{\tyPrimNm{real}}(k) = \{ (x, kx) \sepbar x \in \mathbb{R} \}
  %   \end{array}
  % \end{displaymath}
%   The definitions of $R_{\tyPrimNm{vec}}$ and $R_{\tyPrimNm{real}}$
%   %are as expected. % given the informal definitions we gave in
% formalise the informal ones from~\autoref{sec:motivating-examples}, and 
%   the definitions of %We have set
%   $R^\bullet_{\tyPrimNm{vec}}$ and $R^\bullet_{\tyPrimNm{real}}$ 
%   %to relate $\vec{0}$ and $0$ to themselves to 
%   account for the
%   polymorphic zero constants we have assumed.
%\end{example}

\begin{lemma}\label{lem:geom-environments-1}
  For all $\Delta$ and $\rho \in \relEnv{E}_\Geom(\Delta)$, $(\eta_\Geom,\eta_\Geom) \in \rsem{\Gamma_\Geom}{\relEnv{E}_\Geom}\rho$.
\end{lemma}

\subsection{Free Theorems}
\label{sec:theorems-for-free}

Consider the type of the area %of a triangle 
function from
\exref{ex:area-of-triangle-1}:
\begin{displaymath}
  \begin{array}{@{}l}
    \mathrm{area} : \forall B\mathord:\SynGL{2}, t\mathord:\SynTransl{2}.\ \\
    \hspace{0.8cm}\tyPrim{vec}{B, t} \to \tyPrim{vec}{B, t} \to \tyPrim{vec}{B, t} \to \tyPrim{real}{\abs{\det B}}
  \end{array}
\end{displaymath}
By \thmref{thm:abstraction}, we can derive the following free theorem. For all $B \in \GL{2}$, $\vec{t} \in \Transl{2}$, and $\vec{x}, \vec{y}, \vec{z} \in \mathbb{R}^2$, we have 
\begin{displaymath}
  \abs{\det B}(\tmSem{\mathrm{area}}\ \vec{x}\ \vec{y}\ \vec{z}) = \tmSem{\mathrm{area}}\ (B\vec{x} + \vec{t})\ (B\vec{y} + \vec{t})\ (B\vec{z} + \vec{t})
\end{displaymath}
Thus, directly from the type of the $\mathrm{area}$ function, we can
see that its index-erasure semantics is (a) invariant under
translations, and (b) if the inputs are subjected to a linear
transformation $B$, the output varies with the absolute value of the
determinant of $B$.

@@

\paragraph{Translations} 
Consider first the group $\Transl{2}$ of translations from
\autoref{sec:type-system-geom-intro} and 
\autoref{sec:affine-vector-ops}. For any quantifier-free type $A$
we have: % the isomorphism:
\begin{eqnarray*}
\forall t\mathord:\SynTransl{2}.\tyPrim{vec}{t}\to A
& \cong &
A[0/t]
\end{eqnarray*}
where $A[0/t]$ denotes substitution of the identity translation $0$
for index variable $t$ in type $A$. This isomorphism formalises the
equivalence between \emph{coordinate-free} geometry, expressed by a
polymorphic type whose first argument can be thought of as the
\emph{origin} with respect to which $A$ is specified, and a
\emph{coordinate-based} geometry, expressed by the type $A[0/t]$, in
which the origin is fixed at $(0,0)$. It is sometimes said that an
affine space is a vector space that has forgotten its origin; we have
captured this in a type isomorphism.

\exref{ex:type-iso} is a special case of this isomorphism; another instance is the
type of vector addition:
\begin{eqnarray*}
\forall t_1\mathord:\SynTransl{2}.\forall t_2\mathord:\SynTransl{2}.\tyPrim{vec}{t_1}\to\tyPrim{vec}{t_2}\to\tyPrim{vec}{t_1+t_2}\\
\cong\quad
\forall t_1\mathord:\SynTransl{2}.\tyPrim{vec}{t_1}\to\tyPrim{vec}{t_1}\quad\cong\quad
\tyPrim{vec}{0}
\end{eqnarray*}
\noindent
\newcommand{\transup}[1]{\mathord\uparrow_{#1}}
\newcommand{\transdn}[1]{\mathord\downarrow_{#1}}
The maps associated with the isomorphism are 
$I = \lambda x. x\ [0]\ 0$ and $J = \lambda y. \Lambda t.\lambda v. \transup A (y)$,
%\begin{eqnarray*}
%I = \lambda x. x\ [0]\ 0 & \text{ and } &
%J = \lambda y. \Lambda t.\ \lambda v. \trans A v t(y)
%\end{eqnarray*}
where terms $\transdn A : A\to A[0/t]$ and $\transup A : A[0/t] \to A$ are defined by induction over the syntax of $A$ (cases omitted for $\transup A$ have definition symmetric to $\transdn A$):
\begin{align*}
\transdn{\tyPrimNm{unit}} &= \lambda x. x \qquad
\transdn{A_1\times A_2} = \lambda p. (\transdn{A_1}(\pi_1 p),\transdn{A_2}(\pi_2 p))\\ 
\transdn{A_1 + A_2} &= \lambda x.
\textrm{case}\ x\ \textrm{of}\ \textrm{inl}\ y.\textrm{inl}(\transdn{A_1}y); \textrm{inr}\ y.\textrm{inr}(\transdn{A_2}y)\\ 
\transdn{A_1\to A_2} &= \lambda f. \lambda x. \transdn{A_2} (f(\transup{A_1}x))\\ 
\transdn{\tyPrim{vec}{e}} &= \lambda x. x - kv \qquad
\transup{\tyPrim{vec}{e}} = \lambda x. x + kv\\ 
\text{ if }e & \equiv k t + e' \text{ for
  some }k\in\setOfIntegers \text{ and }e'\text{ with }t\notin e'
\end{align*}
To show $I(J(y))$ is contextually equivalent to $y$ is an easy
induction on $A$ using the fact that the vector $0$
is an identity for vector addition. To show $J(I(x))$ is contextually
equivalent to $x$ we must apply the abstraction theorem and use
the fact that $\tmSem{\transup A}$ and $\tmSem{\transdn A}$ 
considered as relations are simply instances of $\rsem A{}{}$.
%\[
%$\rsem A {t \mapsto v} {} = \{ (x, \tmSem{\transup A (x)}) \mid x \in
%\tySem A\}$.
%\]

@@

\subsection{Non-definability}
\label{sec:types-indexed-abelian-groups-indef}

\newcommand{\Grp}{\mathit{Gr}}

For proving non-definability results, we consider the general case of
a sort $\mathsf{G}$ with the equational theory of abelian groups.  We
assume a group-indexed primitive type $\tyPrimNm{G}$ and set of
primitive operations typed by $\Gamma_\Grp$ that reflect the group
structure at the value level: identity $1:\tyPrim{G}{1}$, operation
$(\cdot): \forall a,b\mathord:\mathsf{G}.\ \tyPrim{G}{a} \to
\tyPrim{G}{b} \to \tyPrim{G}{a\cdot b}$ and inverse $\ ^{-1}: \forall
a\mathord:\mathsf{G}.\ \tyPrim{G}{a} \to \tyPrim{G}{a^{-1}}$. For the
index-erasure semantics, we assume some (not-necessarily abelian)
group $G$ and $\tyPrimSem{\tyPrimNm{G}} = G$.  The interpretation
$\eta_\Grp$ of the primitive operations in $\Gamma_\Grp$ is just the
unit, multiplication and inverse operations of $G$. One instance is the
sort~$\SynTransl{2}$ with $0$, $(+)$ and $\mathrm{negate}$ on
$\tyPrim{vec}{t}$.

We use the construction of non-compositional relational environments
from \autoref{sec:constr-rel-env}, with $R^\bullet_{\tyPrimNm{G}} =
\emptyset$, to give a pre-substitutive family $\relEnv{E}_\Grp$. This
gives us the following lemma and theorem.

\begin{lemma}\label{lem:geom-environments-2}
  For all $\Delta$ and $\rho \in \relEnv{E}_\Grp(\Delta)$,
  $(\eta_\Grp,\eta_\Grp) \in \rsem{\Gamma_\Grp}{\relEnv{E}_\Grp}\rho$.
\end{lemma}

% For index-erasure interpretation, we assume a (not necessarily
% abelian) group $(G, 1, \cdot, \ ^{-1})$. Let $Z(G)$ be the centre of
% the $G$ %: i.e.,~the set of elements that commute with every other
% %element in $G$. We set 
% and $\tyPrimSem{G} = G$. For each index context $\Delta$, the set
% $\relEnv{E}_\Grp(\Delta)$ of relational environments is defined to be
% $\{ \rho_{(H,h)} \sepbar H\textrm{ is a subgroup of }(\Delta
% \Rightarrow \mathsf{G})/\equiv, h\textrm{ is a group homomorphism }H
% \to Z(G) \}$, where
% \begin{displaymath}
%   \rho_{(H,h)}\ \tyPrimNm{G}\ (e) = \left\{
%     \begin{array}{ll}
%       \{ \} & \textrm{if}\ e\not\in H \\
%       \{ (x, h(e)\oplus x) \sepbar x \in G \} & \textrm{if}\ e\in H
%     \end{array}
%   \right.
% \end{displaymath}

Writing $i^n$ for $n$-fold iteration of the group operation, we have:
\begin{theorem}
  Let $m$ and $n$ be natural numbers. For all $n$-by-$m$-matrices of
  natural numbers $A$ and $m$-by-$1$ matricies of integers $b$,
  there exists a quantifier-free program $M$ such that
  \begin{displaymath}
    i_1,\dots,i_m\mathord:\mathsf{G}; \Gamma_\Grp \vdash M : \tyPrim{G}{e_1} \to \dots \to \tyPrim{G}{e_n} \to \tyPrim{G}{e}
  \end{displaymath}
  where $e_j = i_1^{A_{j1}}\cdots i_m^{A_{jm}}$ and $e = i_1^{b_1}\cdots i_m^{b_m}$, iff there is a solution $\vec{x}$
  in the integers to the equation $A \vec x = b$.
\end{theorem}
The uninhabited type $\forall t \mathord: \SynTransl{2}.\
\tyPrim{vec}{t + t} \to \tyPrim{vec}{t}$ from
\exref{ex:uninhabited-type} is an instance of this theorem.  For types
indexed by the scaling group $\SynGL{1}$ the situation is more subtle,
as we have assumed that the type $\tyPrim{real}{s}$ has a polymorphic zero
(see \autoref{fig:real-ops}). We can show, nonetheless, that $\forall
s \mathord: \SynGL{1}.\ \tyPrim{real}{s^2} \to \tyPrim{real}{s}$ is
only inhabited by the constantly zero function by setting
$R^\bullet_{\tyPrimNm{real}} = \{(0,0)\}$. If we augment our sort
$\SynGL{1}$ with a square root operation $s^{1/2}$, then it is
possible to add a primitive square root function and still be able to
prove that the cube root type $\forall s \mathord: \SynGL{1}.\
\tyPrim{real}{s^3} \to \tyPrim{real}{s}$ is only inhabited by the
constantly zero function. This is a consequence of the impossibility
of generating $s$ from $s^3$ with only the abelian group operations
and square root.
